\artigofalse
\chapter{INTRODUÇÃO GERAL}
\shorttitle{Introdução Geral}
\label{appen:introduction-pt}

Modern soil spatial modelling is based on using statistical models to explore the empirical relationship among 
environmental conditions and soil properties. These soil spatial models, like any other model, are nothing 
more than a simplification of reality. Unless we observe the soil everywhere -- which would destroy the 
soil and render the observations useless --, no matter how large the volume of data is, or how comprehensive 
our background knowledge, it will \emph{never} be possible to construct a model that explains the entire 
complexity of the soil. Thus, the outcome of a soil spatial model, i.e. a soil map, will \emph{always} deviate 
from the \q{truth} -- this deviation from the \q{truth} is what we call \emph{error}. What a soil map conveys 
is what we expect the soil to be, acknowledging that there is \emph{uncertainty} about it.

Because soil spatial modellers aim at using the available resources to produce the most accurate 
representation of the soil, a sensible research programme is to investigate the main causes for soil maps 
being more or less \emph{uncertain}. There are many sources of uncertainty in soil spatial modelling, such as 
the errors that result from using a poor statistical model or from making interpolations and extrapolations 
to predict soil properties at unvisited locations. Another important source of uncertainty is the data used to 
assess the empirical relationship among environmental conditions and soil properties: covariate and soil data.

The general objective of this thesis is to evaluate important sources of uncertainty in soil spatial modelling 
with emphasis on soil and covariate data. This general objective can be divided into specific objectives and 
their respective research questions:

\begin{enumerate}[label=(\Roman*)]
 \item Determine the suitability of freely available covariates to calibrate soil spatial models.
  \begin{enumerate}[label=(\alph*)]
   \item Does the use of more detailed covariates result in considerably more accurate soil maps?
   \item How does incorporation of spatial dependence in a soil spatial model compares to the gain in 
   prediction accuracy obtained from using more detailed covariates?
   \item Are the answers to these research questions consistent across soil properties?
  \end{enumerate}
 
 \item Identify the factors that determine how field soil spatial modellers select soil observation locations.
  \begin{enumerate}[label=(\alph*)]
   \item Which factors are considered for deciding upon the location of soil observations, and do they have a 
   pedological origin?
   \item Do the factors play the same role along the course of the soil observation process?
   \item Can point pattern analysis help understanding the purposive sampling strategy traditionally employed
   by field soil spatial modellers?
  \end{enumerate}

\item Identify appropriate calibration sample sizes and designs for soil spatial modelling.
\begin{enumerate}[label=(\alph*)]
\item Can the conditioned Latin hypercube sampling algorithm be improved, and does this improvement deliver 
more accurate soil spatial predictions?
\item Which are the most theoretically sound sampling algorithms for spatial trend estimation, variogram 
estimation, and spatial prediction when we know very little about the soil spatial variation?
\item Can these sampling algorithms be used to construct a general purpose sampling algorithm?

%   \item How suboptimal is it to use a sample configuration that was optimized for a different purpose?
%   \item Is the predictive performance of a soil-mapping model estimated using a sample configuration 
%   optimized using heuristics poorer than that of another soil-mapping model whose parameters were estimated 
%   using a sample configuration optimized using an \textit{a priori} knowledge of the model?
%   \item Is it possible to obtain a sample configuration that is efficient in identifying and estimating i) 
%   the spatial trend and ii) the variogram model, and iii) making spatial predictions?
%   \item How does the sample configuration affect the estimated model parameters and thus the conclusions 
% that 
%   can be drawn under the light of the existing conceptual model of pedogenesis?
%   \item Are the answers to the research questions above consistent across sample sizes and soil properties?
  \end{enumerate}
\end{enumerate}

The thesis is composed of eight chapters where each of the above mentioned objectives are met. Although there 
is a logical sequence in their presentation, all chapters were planned so that they could be read separately. 
This means that there is some overlap between them, i.e. repeated information. References to specific sections 
of other chapters using coloured (blue) hyperlinks are common.

\autoref{chap:chap02} is a commented review of the literature on soil spatial modelling and its main sources 
of 
uncertainty. The review starts with a discussion about the efforts made by soil spatial modellers to raise 
awareness about the importance of soil spatial information. These efforts seem to have fuelled a global 
scientific demand for up-to-date, high resolution soil spatial information. The chapter continues with a 
description of soil spatial modelling along human history, suggesting that the goal of producing soil maps 
remains more or less the same since the Neolithic Revolution (ca.~\num{10000}~years). The chapter closes with 
the main sources of uncertainty.

\autoref{chap:chap03} presents the conceptual model of pedogenesis (in Portuguese), which consists 
of a description of the study area that includes an explicit description of soil-forming factors (climate, 
geology, geomorphology, hydrology, land use, and vegetation) and processes that determine the soil 
spatio-temporal distribution. \autoref{chap:chap04} describes the soil data included in the 
\emph{Santa Maria dataset}, which was used to develop the case studies presented in this thesis. The Santa
Maria dataset is composed of $n = 410$ soil observations compiled from studies carried out between \num{2004} 
and \num{2013}. These studies aimed at producing semi-detailed soil and land use maps, and modelling topsoil 
carbon stock and vulnerability to erosion. A comprehensive description of the covariate data included in the 
Santa Maria dataset, and their processing, is given in \autoref{chap:chap05}. Beyond describing 
the data used in the thesis, the goal of these two chapters, along with the conceptual model of pedogenesis, 
is 
to provide the basis for future soil spatial modelling exercises in the study area, and to serve as examples 
for new soil spatial modelling studies developed elsewhere.

Based on an article published in the peer reviewed journal \geoderma, \autoref{chap:chap06} serves 
the purpose of meeting the first objective of the thesis and answering its respective research questions. The 
prediction performance of linear soil spatial models calibrated using covariates (area-class soil maps, land 
use maps, geological maps, digital elevation models, and satellite images) available in two levels of detail 
is evaluated. The influence of taking the spatial dependence of the residuals into account is assessed as 
well. 

\autoref{chap:chap07} presents an approach that aims at helping to understand the purposive 
sampling strategy traditionally employed by field soil modellers, i.e. free survey. This is important because 
many soil spatial modelling projects rely on legacy data, i.e. soil data produced previously and made 
available 
(publicly or not), whose observation locations were purposively selected by soil spatial modellers using 
poorly documented tacit rules. The chapter is designed to answer the research questions of the second 
objective 
of the thesis. Point pattern analysis is used to characterize the spatial sample configuration, 
whereas theories borrowed from Psychology are used to elaborate on the subjective factors involved in 
selecting 
soil observation locations.

Objective 3 and its research questions are addressed in \autoref{chap:chap08} and \autoref{chap:chap09}. In 
\autoref{chap:chap08}, three improved sampling algorithms are compared to the original conditioned Latin 
hypercube sampling algorithm on how they affect geographic coverage, estimated model parameters and prediction 
accuracy. The influence of sample size is also discussed. \autoref{chap:chap09} presents the most efficient 
sampling strategies for spatial trend estimation, variogram estimation, and spatial prediction when we know 
very little about the soil spatial variation. The chapter closes with a new general purpose sampling algorithm 
that aims at the three objectives jointly.

The sequence of eight chapters is closed with \hyperref[chap:chap10]{General Conclusions} where I highlight 
the main research results and contributions of the study. Next, there are two appendices, both devoted to the 
description of the two packages for \texttt{R} developed to support the thesis: \texttt{spsann} 
(\autoref{appen:spsann}) and \texttt{pedometrics} (\autoref{appen:pedometrics}). The first was designed for 
the optimization of sample configurations using spatial simulated annealing. The second includes miscellaneous 
functions that were put together for ease of use. All literature references are presented under a unique list 
of \hyperref[chap:references]{Bibliographic References} at the end of the thesis.

